\Large{ Authors' reply to the comments by M. deâ Michieli Vitturi (Referee \#1) of the manuscript gmd-2017-119} \\ \LARGE \textbf{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics}

### 2017-09-21 Reply-to-RC1.tex
9/20 edits ap

\documentclass[10pt,a4paper]{article}
%\documentclass[10pt,a4paper]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{kpfonts}
\usepackage[numbers]{natbib}
%other options include: round, curly, angle, or default options

\usepackage{media9}

\author{\Large \textbf{by Zhixuan Cao et al.}}

\title{ \Large{ Authors' reply to the comments by M. de’ Michieli Vitturi (Referee \#1) of the manuscript gmd-2017-119} \\
\LARGE \textbf{“Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics”}}

@@

\begin{document}
\maketitle
The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers that we refered in our response are corresponding to these in the original varsion of GMD discussion paper. \\[12pt]

@@

\textit{This paper presents an initial effort towards developing a plume model with comprehensive physics based on the smoothed particles hydrodynamics method. To my knowledge, this is the first numerical model of volcanic plumes adopting this technique and this is a good addition to the existing models. While the model, the derivation of the discretised equations and the computational techniques are satisfactory for a reader expert in SPH, a few more details would help in the comprehension other readers. The scientific content of the paper could be improved by a better and extended description of the applications. Overall, the manuscript represent a substantial contribution to modelling Âascience within the scope of Geoscientific Model Development and it is suitable for publication after some improvements and corrections are made. These are described below.}

@@

The following are our response to reviewer's specific comments. Modifications are made accordingly in the manuscript. \\[12pt]

@@

Yes, dynamic equilibrium means same velocity while thermodynamic equilibrium means same temperature. We treat erupted material (includes volcanic gas and particles) as one phase, so assumption of dynamic equilibrium and thermodynamic equilibrium are made for erupted material. Such assumption is made implicitly when we assuming erupted material behaves like a single phase fluid.
What we intended to say by "dynamic and thermodynamic equilibrium between air and erupted material" is actually "dynamic equilibrium and thermodynamic equilibrium between erupted material and air that is entrained into the plume". That is to say, we assume that surrounding air and erupted material will reach an equilibrium status (same velocity and temperature) as soon as they get mixed.

We are aware of that the original statement is misleading and modified it as "dynamic equilibrium and thermodynamic equilibrium between erupted material and air that is entrained into the plume". We thanks reviewer for pointing this out. \\[3pt]

@@

Thank you for your clarification on this. We revised original statement "The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991." to "The code is first verified by 1D shock tube tests, then by comparing velocity and concentration distribution along the central axis and on the transverse
cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment). 
%The model is validated by comparing against simulation results from other 3D plume models.
Profiles of several integrated variables are compared with those calculated by existing 3D plume models for an eruption with the same MER (mass eruption rate) estimated for the Pinatubo eruption of June 15 1991. Our results are consistent with existing 3D plume models." \\[6pt]

@@

\textit{Page 4, lines 23-25. It is not clear to me why interface tracking or interface capturing are mentioned here for mesh based method. I think all the existing 3D Eulerian models (based on mesh based method) do not model the interface between the plume and the
atmosphere. While I understand that this represent a big advantage for other applications of SPH, for example for dam-break problems, because it allows to solve only for the “flow” region and not for the surrounding, for a volcanic plume mixing between plume and air is important and I don’t really see a clear interface between them. Please clarify and expand the point.}

We first thank reviewer for asking this question. This question pushed us to do more investigation on "interface construction" in mesh based methods. More importantly, we got an deeper understanding on why we need turbulence model in plume modeling. Our original motivation for using a turbulence model was simply based on the following logic: The Reynolds number is large enough hence the flow is turbulent, we can not resolve all turbulent exchange at all different scales with a coarse resolution, so we need turbulence model to resolve sub-particle scale turbulence. Now we can see the necessary of including a turbulence model in purpose of resolving mixing which happens in a much smaller scale.

Because interface acts as portion of boundaries.
It is critical to capturing or tracking the interface during simulation for immiscible flow. This is one of major motivations for using interface tracking and interface capturing method. However, this is not the case for volcanic plume modeling. As "fluids" in our plume model are miscible.

Even though, clear transient interface do exist in miscible flow. There are experiments \citep {papantoniou1989large} showing clear interface between air and plume. Another experiment with higher resolution showing more detailed and more complicated interface by \citet{crimaldi2001high}. 
It is through the interface (or boundary) where mxing happens. \citet{jacobson2008mixing} study detailed mixing mechanism at a flat interface.

However, quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large–scale fluid motion and the diffusion processes on interface that ultimately lead to mixing. The energy transfer between these scales occurs through turbulent motion, created either by fluid instabilities or by breaking internal waves. For shearing flow, like volcanic plume, the fluid instabilities dominated the mixing process. Ideally, one would like to be able to include the effects of mixing on the large scale dynamics without resolving the detailed interface structure and dynamics of turbulence to reduce computational cost. Such strategy was also used in all other mesh-based 3D plume models (though different turbulence closures were adopted by different models). 
By resolving sub-particle scale (or sub-mesh scale) turbulent mixing with turbulence closure, relative coarse resolution is used in all 3D plume models (both particle based model and mesh-based models). Such solution is a trade-off between computational accuracy and computational cost. And in this case, interface might not be clearly observed in simulation due to coarse resolution.

@@

Remedies of standard SPH is needed if detailed mixing process needs to be resolved, for example \citep{cha2010kelvin, zhuimproved}.
Remedies, such as adding of surface tension \citep{behjati2017modelling}, are needed for tracking interface while simulating immiscible flow using SPH. Even though, as a Lagrangian method, interface constructing is explicit through capturing of the locations of the particles and much simplier than Eulerian Methods.

Modification in the manuscripts:
1) Since the exact location of interface is less important for plume modeling when turbulence model is adopted, we changed this paragraph to: 
"Interface tracking is explicit in SPH through capturing of the locations of the particles. Less numerical effort is required for interface construction when we attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence." and put it as the last item among SPH features.
2) Added a short paragraph (Basically the fourth paragraph of this response) at the beginning of section 3.8 explaining motivation of using turbulence model. 
3) Changing the title of section 3.7 to "Mass fraction update". Re-organize this section and focus more on mass fraction updating. Make a brief remark on interface constructing using SPH and mesh-based methods at the end of this section. \\[6pt]

@@

\textit{Page 12, line 22. The “sink” term has been added to eq. 42, representing the discretised form of mixture equation. If the term is associated with a phase change, it would be better add it in the equation for a phase and not for the mixture with density $\rho$.}

First of all we would like to calrify that the purpose of showing Eq. (42) ~ (44) is to show that only slight modifications is needed in discretized governing equations when new physics is considered. In addition, we do not have separate equation for mass conservation of volcanic gas (due to the assumption that erupted material behaves like a single phase). Appologize if reviewer has already been aware of this.

We completely agree with the reviewer that for “sink” associated with phase change should be in the equation of the corresponding phase. For any future extending of current model, such as the example mentioned by the reviewer, considering phase change of volcanic gas, such sink term should be added to mass conservation equation of volcanic gas. The make the article more readable, we prefer to avoiding discussing new physics models, which is easily understandable for readers working on plume modeling but might cause confusion for readers without such background.
%To avoid confusing readers, we emphasized at the end of section 3.3:  “Eq. (42) - Eq. (44) show examples of adding more physics into SPH model based on Euler equations to show easy extensibility of SPH models. The mass source (and sink), drag force effect, and heat exchange effect are not actually added to our model (Eq. (1) - (4)) "
\\[3pt]

\textit {Page 12, lines 24-25. “ The drag force term should show up only when dynamics disequilibrium between different phases is considered”. In this case is the summation in the last term of eq. 43 only extended to particles of a different phase?}

@@

If such disequilibrium need to be considered in future extension of current model, drag force effect should be considered. How many drag force effects should be considered depends on new assumptions and the new physics model. For example, if the physics model is a four phases model, one phase for atmosphere air, one phase for volcanic gas, one phase for large size solid particles and another for small size ash (maybe immediate equilibrium assumption between air and gas should be made). Drag effects between gas (air) and particles should be considered. But there should be no drag force term within gas phase itself. As for the small size ash, we can assume them behaves like single phase fluid (possibly not a reasonable assumption) then there should not drag force effect within the ash phase. Otherwise. drag effect within fine ash phase need to be considered.

We changed the original sentence “The drag force term should show up only when dynamics disequilibrium between different phases is considered” to “The drag force term should show up when dynamics disequilibrium between different phases is considered”. By deleting the word “only”, we change original necessary and sufficient statement to a sufficient statement. \\[6pt]

@@

\textit{Page 13, line 5. Please define “particle disorder”.}

We replaced “particle disorder” with “irregular distribution of particles” \\[6pt]

@@

It is calculated by $c=(gamma_m* \frac{p}{/rho•} )^0.5$, we added an equation in our manuscript.  \\[3pt]

@@

I am supposing that you are asking about how much is each time step.

@@

\textit{Page 14, line 8. “The classical SPH method was known to suffer from tensile instability and boundary deficiency.” Please describe there problems with some example, in order
to help the readers not expert in SPH.}

We added explanations and references on tensile instability: "Tests of the standard SPH method indicate an instability in the tensile regime, while the calculations are stable in compression.  A simplest example of such test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile. If the initial stress is tensile, a very small velocity perturbation on a single particle can lead to particles clumping together, forming large voids and seriously corrupting density distribution. But if the initial stress is compressive, the small velocity perturbation on a single particle can not lead to any changes in particle distribution. See paper by \citet{swegle1995smoothed} for more details." \\[3pt]

\textit{Page 14, line 14. “Equation (55) implies that $. . ..=1$”. Why? It is not equation 55 which implies the term is equal to 1, but the properties of the kernel. Furthermore, this is
true for the integral (not discretised) formulation, and only for points far away from the boundary (see Chen et al., 1999).}

The sentence is a misleading expression. Sorry for that. Here is the logic behind: The new approximation of function $A$, Eq. (55), should be consistent with original formulation, Eq. (28). Compare Eq. (55) and Eq. (28), if the denominator is 1, then these two equations will be the same. so it implies that the denominator in Eq. (55) should be close to 1. 
We completely agree with you that the normalization condition originally come from definition of kernel. By the statement "implies that $...=1$" we were actually trying to show kind of consistent connection between this new approximation formulation and property of kernel function. In other words, in the derivation based on Taylor series expansion, we never take the property of kernel into account but finally reach to an equation that consistent with one property of kernel function.

@@

We changed the original sentence "Equation (55) implies that $\sum_b	m_b \rho_b w (x-x_b, h) = 1$ , which can be viewed as the approximation form of Eq. (34)." to  "Notice that the denominator in Eq. (55) is actually summation approximation of Eq. (34). That is to say, Eq. (55) and Eq. (28) are the same for particles far away from boundaries as the denominator in Eq. (55) becomes $1$ in that case." \\[6pt]

@@

\textit{Page 14, line 21-22. “Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.”
This is true when the different phases are immiscible. But in the application investigated, the volcanic plume, phases are not immiscible and mixing is very important. So, there is no need to track the interface between phases Eulerian grid-based numerical methods. Conversely, standard formulations of SPH cannot resolve fluid mixing and instabilities at flow boundaries. IT should be discussed in the paper if, and how, this is has been addressed in the model presented. Please take a look at: J. I. Read, T. Hayfield, O. Agertz; Resolving mixing in smoothed particle hydrodynamics, Monthly Notices of the Royal Astronomical Society, Volume 405, Issue 3, 1 July 2010, Pages 1513–1530}

@@

The following is our response to reviewer's concern on how did we handle the issue of classical SPH: "Classical SPH has problems correctly integrating fluid instabilities and mixing at boundaries". I will call it "mixing challenge" in later paragraph for short.

We were not aware of Read's paper previously. But we adopted several techniques (corrected formulation for tensile instability issue, smaller artificial viscosity coefficients, turbulence model with heat transfer considered) that probably have similar affect as the methods proposed in Read's paper. In Read's paper, one of his remedies of SPH is based on fixing the clumping instability which is actually tensile instability. We just adopted a different method (Chen's method) to handle this issue.

In addition, before we adopting the turbulence model, we got a "mushroom like" plume and almost no mixing happened (exactly due to the "mixing challenge"). The simulation results became much more realistic after adding turbulence model. So the turbulence model helped us to resolve sub-particle scale mixing. As has been mentioned, it is a common practice in CFD to adopt turbulence closure to resolve instabilities (which dominate the mixing process for shearing flow).

@@

\citet{price2008modelling} believes that the mixing issue is due to the fact that entropy is discontinuous at the boundaries while density is continuous. He found that adding thermal conductivity at boundaries can improve mixing in SPH as thermal conductivity can smooth the entropy. The thermal conductivity (heat transfer) due to turbulence is considered in our model. What's more the thermal conductivity coefficient is much larger at the interfaces due to large shearing effect at the interface. Which can definitely help to releive the "mixing challenge" if Price is correct. Of course, we were not motivated by Price's paper to adding a turbulent heat transfer term into our model. It is natural to have a turbulent heat transfer term for compressible flow in which energy conservation equation is coupled with momentum conservation and mass conservation equation. As you can see we actually and luckly "adopted" Price's method.

\citet{wadsley2008treatment} and \citet{ritchie2001multiphase} made similar argument as Price. \citet{borgani2012hydrodynamic} believes that the "mixing challenge" of SPH is due to its poor ability of capturing contact discontinuity. They show that using GSPH can avoid appearance of spurious pressure force and help to to follow the kelvin instability.

Since people still have different opinions on the sources of "mixing issue", it would be interesting to see the connections between these different opinions. But such work is obviously out of our topic in this paper. These remedies of traditional SPH we adopted and extended in our model helped us to overcome, at least relieve the "mixing challenge".

@@

\textit{Page 16, line 10. For a reader like me who is more familiar with Eulerian and meshbased formulation, it would be good to state at the beginning that, differently from RANS (Reynolds Averaged Navier-Stokes) equations, which are time-averaged equations of motion, here the Lagrangian average is in space and not in time.}

Thank you for your suggestion for improving readability of the article. We totally agree with you and have made revision as you suggested. \\[6pt]

@@

In our current model, the ground is assumed to be a flat ground. Since the topography is pretty simple, we did not experience any trouble. it has been shown in some other papers that it works as well for more complex topography \citep{kumar2013parallel}. One potential issue for complicated topography might be deployment of wall ghost particles. For flat boundary, we simply deploy several layers of ghost particles with equal interval and each ghost particle has the same mass. It has not been investigated yet how different deployment of wall ghost particle will influence the accuracy of imposed boundary conditions. A more basic question is how many degree of freedom should we allow for ghost particle deployment: only allows changing of particle position? Or allows using different particle mass size?

For non-collapse plume, the flat ground assumption made in our current model should be good enough. For other phenomena, such as PDCs, a more realistic topography is required. As we mentioned in the manuscripts, current plume models are still not comprehensive enough, each model has its own problem of interest and assumptions were made accordingly. Fortunately, we’ve been seeing efforts on making these models to be more general. As the first trial of using SPH in plume modeling, we prefer to narrowing down the problem of interesting and focus on more fundamental aspects.

@@

I have two questions regarding your suggestion: 
1) When we assign pressure value to these static pressure ghost particles, the decrease of pressure due to increase of height ($h$) has already been took into account. That is to say, pressure of our pressure ghost particles is consistent with meteorological data when establishing the initial condition. I am wondering should I still deduce $\rho g h$ from pressure?  
2) Should the total pressure equals to a pressure \textbf{plus} kinetic energy, or minus kinetic energy? I have similar question for definition of $p_{\rho}$. Should $\rho g h$ been added to or deduced from pressure? Because total pressure in other fields, such as thermal dynamic analysis of flowing process in engines are definied as pressure plus kinetic energy. Just to make sure.

In our current model, pressure ghost particles are assumed to be stationary. It does not make difference whether we use pressure or total pressure since velocity of pressure ghost particles are zero. The next important improvement of our current model under planning is taking the effect of wind field into account. In that case, the pressure ghost particles will not be static any more. The idea of using total pressure can be naturally implemented for that case. Thank you very much for pointing out an potential way for improving pressure outlet boundary condition. \\[6pt]

@@

\textit{deals with the numerical resolution of the equations of the model, not with the agreement between the model and reality. It checks no code errors have been introduced
in the code and one way to do it is a comparison with analytical solutions. I think it is better to simply rename the section "Applications” or “Results”.}

Since JPUE is not the phenomena that we are targeted to model. So that is not validation (validation is to make sure assumptions in physics model is consistent with reality). Meanwhile, we compared our results against experimental results. We might call such comparison (simulation against measurements) as verification in a more general sense (?).

We added simulation results of shock tube tests which were compared against analytic solutions. By adding such real verification tests, we keep the title of this section unchanged.
\\[6pt]

@@

\textit{Page 25, line 24. Please look also at the results from this paper: Ezzamel, Adam, Pietro Salizzoni, and Gary R. Hunt. "Dynamical variability of axisymmetric
buoyant plumes." Journal of Fluid Mechanics 765 (2015): 576-611.}

Thank you for pointing out a source for accessing more recent experimental results. We’ve added the paper in the reference. \citet{ezzamel2015dynamical} did not provide coefficients of their fit equations. So we did not add plots corresponding to  Ezzamel's results in our figures. \\[3pt]

>>>>

\documentclass[10pt,a4paper]{article}
%\documentclass[10pt,a4paper]{beamer}
%\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}
\usepackage{kpfonts}
\usepackage[numbers]{natbib}
%other options include: round, curly, angle, or default options

\usepackage{media9}
\begin{document}
\author{\Large \textbf{by Zhixuan Cao et al.}}

\title{ \Large{ Authors' reply to the comments by M. deâ Michieli Vitturi (Referee \#1) of the manuscript gmd-2017-119} \\
\LARGE \textbf{Plume-SPH 1.0: A three-dimensional, dusty-gas volcanic plume model based on smoothed particle hydrodynamics}}

@@

\maketitle
The comments by the reviewer are recited in italics, followed by our reply in upright font. Equation, section, and figure numbers that  in our response are corresponding to these in the original 
version of GMD discussion paper. \\[12pt]

@@

\textit{This paper presents an initial effort towards developing a plume model with comprehensive physics based on the smoothed particles hydrodynamics method. To my knowledge, this is the first numerical model of volcanic plumes adopting this technique and this is a good addition to the existing models. While the model, the derivation of the discretized equations and the computational techniques are satisfactory for a reader expert in SPH, a few more details would help in the comprehension other readers. The scientific content of the paper could be improved by a better and extended description of the applications. Overall, the manuscript represent a substantial contribution to modelling science within the scope of Geoscientific Model Development and it is suitable for publication after some improvements and corrections are made. These are described below.}

@@

The following are our responses to reviewer's specific comments. Modifications are made accordingly in the manuscript. \\[12pt]

@@

Yes, dynamic equilibrium means same velocity while thermodynamic equilibrium means same temperature. We treat erupted material (includes volcanic gas and particles) as one phase, so assumption of dynamic equilibrium and thermodynamic equilibrium are made for erupted material. Such assumption is made implicitly when we assume that erupted material behaves like a single phase fluid.
What we intended to say by ``dynamic and thermodynamic equilibrium between air and erupted material" is actually ``dynamic equilibrium and thermodynamic equilibrium between erupted material and air that is entrained into the plume". That is to say, we assume that surrounding air and erupted material will reach an equilibrium status (same velocity and temperature) as soon as they get mixed.

We realize that the original statement is misleading and modified it as ``dynamic equilibrium and thermodynamic equilibrium between erupted material and air that is entrained into the plume". We thank the reviewer for pointing this out. \\[3pt]

@@

Thank you for your clarification on this. We revised our original statement ``The model is verified by comparing velocity and concentration distribution along the central axis and on the transverse cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment) and the top height of the Pinatubo eruption of 15 June 1991." to ``The code is first verified by 1D shock tube tests, then by comparing velocity and concentration distribution along the central axis and on the transverse
cross with experimental results of JPUE (jet or plume that is ejected from a nozzle into a uniform environment). 
%The model is validated by comparing against simulation results from other 3D plume models.
Profiles of several integrated variables are compared with those calculated by existing 3D plume models for an eruption with the same MER (mass eruption rate) estimated for the Pinatubo eruption of June 15 1991. Our results are consistent with existing 3D plume models." \\[6pt]

@@

\textit{Page 4, lines 23-25. It is not clear to me why interface tracking or interface capturing are mentioned here for mesh based method. I think all the existing 3D Eulerian models (based on mesh based method) do not model the interface between the plume and the
atmosphere. While I understand that this represent a big advantage for other applications of SPH, for example for dam-break problems, because it allows to solve only for the flow region and not for the surrounding, for a volcanic plume mixing between plume and air is important and I don't really see a clear interface between them. Please clarify and expand the point.}

We first thank the reviewer for asking this question. This question pushed us to do more investigation on ``interface construction" in mesh based methods. 
%Because interface acts as portion of boundaries.
It is critical to capture or track the interface during simulation for immiscible flow since the interfaces are the boundaries separating different fluids. This is the major motivation for using interface tracking and interface capturing method in multi-phase flows. However, this is not the case for volcanic plume modeling as ``fluids" in our plume model are miscible.

However, the literature reveals that clear transient interfaces do exist in miscible flow. There are experiments \citep {papantoniou1989large} showing a sharp interface between air and plume. \citet{crimaldi2001high} present another experiment with higher resolution that shows an even more detailed and more complicated interface. We postulate that it is  
through the interface (or boundary) that mixing happens. Analogously, \citet{jacobson2008mixing} study has detailed mixing mechanism at a flat interface.

However, quantifying these mixing processes in real implementation is challenging because of the scale disparity between the large-scale fluid motion and the diffusion processes on interface that ultimately lead to mixing. The energy transfer between these scales occurs through turbulent motion, created either by fluid instabilities or by breaking internal waves. For shearing flow, like volcanic plume, the fluid instabilities dominated the mixing process. Ideally, one would like to be able to include the effects of mixing on the large scale dynamics without resolving the detailed interface structure and dynamics of turbulence to reduce computational cost. Such a strategy was also used in all other mesh-based 3D plume models (though different turbulence closures were adopted by different models).

More importantly, thanks to this reexamination of our approach we now have a deeper understanding of why we need a turbulence model in plume modeling. Our original motivation for using a turbulence model was simply based on the following logic: The Reynolds number is large enough hence the flow is turbulent, we can not resolve all turbulent exchange at all different scales with a coarse resolution, so we need turbulence model to resolve sub-particle scale turbulence. Now we can see the necessity of including a turbulence model to resolve mixing which happens at a much smaller scale.

By resolving sub-particle scale (or sub-mesh scale) turbulent mixing with turbulence closure, relative coarse resolution is used in all 3D plume models (both particle based model and mesh-based models). Such solution is a trade-off between computational accuracy and computational cost. And in this case, interface might not be clearly observed in simulation due to coarse resolution.

@@

Remedies of standard SPH is needed if detailed mixing processes need to be resolved, for example \citep{cha2010kelvin, zhuimproved}.
Remedies, such as adding of surface tension \citep{behjati2017modelling}, are needed for tracking interface while simulating immiscible flow using SPH. Even though, as a Lagrangian method, interface constructing is explicit through capturing of the locations of the particles and much simpler than Eulerian Methods.

Modification in the manuscripts:
1) Since the exact location of interface is less important for plume modeling when turbulence model is adopted, we changed this paragraph to: 
``Interface tracking is explicit in SPH through capturing of the locations of the particles. Less numerical effort is required for interface construction when we attempt to include the effects of mixing by resolving the detailed interface structure and dynamics of turbulence." and put it as the last item among SPH features.
2) Added a short paragraph (Basically the paragraph above of this response) at the beginning of section 3.8 explaining the motivation for using a turbulence model. 
3) Changing the title of section 3.7 to ``Mass fraction update" we  re-organize this section and focus it more on mass fraction updating. We make a brief remark on interface construction using SPH and mesh-based methods at the end of this section. \\[6pt]

@@

\textit{Page 12, line 22. The sink term has been added to eq. 42, representing the discretized form of mixture equation. If the term is associated with a phase change, it would be better to add it in the equation for a phase and not for the mixture with density $\rho$.}

First of all we would like to clarify that the purpose of showing Eq. (42) ~ (44) is to show that only slight modifications is needed in discretized governing equations when new physics is considered. In addition, we do not have separate equation for mass conservation of volcanic gas (due to the assumption that erupted material behaves like a single phase). %Apologize if reviewer has already been aware of this.

We completely agree with the reviewer that terms for sink associated with phase change should be in the equation of the corresponding phase. For any future extensions of current model, such as the example mentioned by the reviewer, considering phase change of volcanic gas, such sink term should be added to mass conservation equation of volcanic gas. To make the article more readable, we avoid
 discussing new physics models, which are easily understandable for readers working on plume modeling but might cause confusion for readers without such background.
%To avoid confusing readers, we emphasized at the end of section 3.3:  Eq. (42) - Eq. (44) show examples of adding more physics into SPH model based on Euler equations to show easy extensibility of SPH models. The mass source (and sink), drag force effect, and heat exchange effect are not actually added to our model (Eq. (1) - (4)) "
\\[3pt]

\textit {Page 12, lines 24-25. The drag force term should show up only when dynamics disequilibrium between different phases is considered. In this case is the summation in the last term of eq. 43 only extended to particles of a different phase?}

@@

If such disequilibrium needs to be considered in future extension of current model, drag force effect should be considered. The form used to represent drag force effects will depend on new assumptions and the new physics model. For example, if the physics model is a four phases model, one phase for atmosphere air, one phase for volcanic gas, one phase for large size solid particles and another for small size ash (maybe immediate equilibrium assumption between air and gas should be made). Drag effects between gas (air) and particles should be considered. But there should be no drag force term within gas phase itself. As for the small size ash, we can assume they behave like single phase fluid (possibly not a reasonable assumption) then there should not drag force effect within the ash phase. Otherwise. drag effect within fine ash phase need to be considered.

We changed the original sentence ``The drag force term should show up only when dynamics disequilibrium between different phases is considered" to ``The drag force term should show up when dynamics disequilibrium between different phases is considered. By deleting the word only we changed original necessary and sufficient statement to a sufficient statement. \\[6pt]

@@

\textit{Page 13, line 5. Please define particle disorder.}

We replaced particle disorder with irregular distribution of particles \\[6pt]

@@

It is calculated by $c=(\gamma_m* \frac{p}{\rho} )^{0.5}$ -- we added an equation in our manuscript.  \\[3pt]

@@

%I am supposing that you are asking about how much is each time step.

@@

\textit{Page 14, line 8. The classical SPH method was known to suffer from tensile instability and boundary deficiency. Please describe there problems with some example, in order
to help the readers not expert in SPH.}

We added explanations and references on tensile instability: ``Tests of the standard SPH method indicate an instability in the tensile regime, while the calculations are stable in compression.  A simple test calculation exhibiting the instability involves a body which is subjected to an uniform initial stress, either compressive or tensile. If the initial stress is tensile, a very small velocity perturbation on a single particle can lead to particles clumping together, forming large voids and seriously corrupting density distribution. But if the initial stress is compressive, the small velocity perturbation on a single particle can not lead to any changes in particle distribution. See paper by \citet{swegle1995smoothed} for more details." \\[3pt]

\textit{Page 14, line 14. Equation (55) implies that $. . ..=1$â Why? It is not equation 55 which implies the term is equal to 1, but the properties of the kernel. Furthermore, this is
true for the integral (not discretised) formulation, and only for points far away from the boundary (see Chen et al., 1999).}

The sentence is misleading. Sorry for that. Here is the logic: The new approximation of function $A$, Eq. (55), should be consistent with original formulation, Eq. (28). Compare Eq. (55) and Eq. (28), if the denominator is 1, then these two equations will be the same. so it implies that the denominator in Eq. (55) should be close to 1. 
We completely agree with you that the normalization condition originally come from definition of kernel. By the statement ``implies that $...=1$" we were actually trying to show a  consistent connection between this new approximation formulation and properties of the kernel function. In other words, in the derivation based on Taylor series expansion, we never take the property of the kernel into account but finally reach to an equation that is consistent with a property of kernel function.

@@

We changed the original sentence ``Equation (55) implies that $\sum_b	m_b \rho_b w (x-x_b, h) = 1$ , which can be viewed as the approximation form of Eq. (34)." to  ``Notice that the denominator in Eq. (55) is actually summation approximation of Eq. (34). That is to say, Eq. (55) and Eq. (28) are the same for particles far away from boundaries as the denominator in Eq. (55) becomes $1$ in that case." \\[6pt]

@@

\textit{Page 14, line 21-22. Numerical simulation of multiphase flows is usually difficult due to the existence of complex evolving interfaces between phases.â
This is true when the different phases are immiscible. But in the application investigated, the volcanic plume, phases are not immiscible and mixing is very important. So, there is no need to track the interface between phases as in Eulerian grid-based numerical methods. Conversely, standard formulations of SPH cannot resolve fluid mixing and instabilities at flow boundaries. IT should be discussed in the paper if, and how, this is has been addressed in the model presented. Please take a look at: J. I. Read, T. Hayfield, O. Agertz; Resolving mixing in smoothed particle hydrodynamics, Monthly Notices of the Royal Astronomical Society, Volume 405, Issue 3, 1 July 2010, Pages 15131530}

@@

The following is our response to reviewer's concern on how did we handle the issue of classical SPH: ``Classical SPH has problems correctly integrating fluid instabilities and mixing at boundaries". We will call it ``mixing challenge" in later paragraph for short.

We were not aware of Read's paper previously. But we adopted several techniques (corrected formulation for tensile instability issue, smaller artificial viscosity coefficients, turbulence model with heat transfer considered) that probably have similar effect as the methods proposed in Read's paper. In Read's paper, one of his remedies of SPH is based on fixing the clumping instability which is actually tensile instability. We just adopted a different method (Chen's method) to handle this issue.

In addition, before we integrated the turbulence model, we got a ``mushroom like" plume and almost no mixing happened (exactly due to the ``mixing challenge"). The simulation results became much more realistic after adding the turbulence model. So the turbulence model helped us to resolve sub-particle scale mixing. As has been mentioned, it is a common practice in CFD to adopt turbulence closure to resolve instabilities (which dominate the mixing process for shearing flow).

@@

\citet{price2008modelling} believes that the mixing issue is due to the fact that entropy is discontinuous at the boundaries while density is continuous. He found that adding thermal conductivity at boundaries can improve mixing in SPH as thermal conductivity can smooth the entropy. The thermal conductivity (heat transfer) due to turbulence is considered in our model,  what is more is that the thermal conductivity coefficient is much larger at the interfaces due to large shearing effect at the interface. This can definitely help to relieve the ``mixing challenge" if Price is correct. While, we were not motivated by Price's paper to adding a turbulent heat transfer term into our model -- we were motivatd by the need to have a turbulent heat transfer term for compressible flow in which energy conservation equation is coupled with momentum conservation and mass conservation equation. It appears we were lucky and ``adopted" Price's method.
\citet{wadsley2008treatment} and \citet{ritchie2001multiphase} made similar arguments as Price. \citet{borgani2012hydrodynamic} believes that the ``mixing challenge" of SPH is due to its poor ability of capturing contact discontinuity. They show that using GSPH can avoid appearance of spurious pressure force and help to to follow the Kelvin instability.
Since people still have different opinions on the sources of ``mixing issue", it would be interesting to see the connections between these different opinions. But such work is obviously out of our scope in this paper. These remedies of traditional SPH we adopted and extended in our model helped us to overcome or at least relieve the ``mixing challenge".

@@

\textit{Page 16, line 10. For a reader like me who is more familiar with Eulerian and mesh based formulation, it would be good to state at the beginning that, differently from RANS (Reynolds Averaged Navier-Stokes) equations, which are time-averaged equations of motion, here the Lagrangian average is in space and not in time.}

Thank you for your suggestion for improving readability of the article. We totally agree with you and have made the revision as you suggested. \\[6pt]

@@

In our current model, the ground is assumed to be a flat ground. Since the topography is pretty simple, we did not experience any trouble. it has been shown in some other papers that it works as well for more complex topography \citep{kumar2013parallel}. One potential issue for complicated topography might be deployment of wall ghost particles. For flat boundary, we simply deploy several layers of ghost particles with equal interval and each ghost particle has the same mass. It has not been investigated yet how different deployment of wall ghost particle will influence the accuracy of imposed boundary conditions. A more basic question is how many degree of freedom should we allow for ghost particle deployment: only enough to allow changing of particle position  or enough to allow using different particle mass size?

For non-collapsed plume, the flat ground assumption made in our current model should be good enough. For other phenomena, such as PDCs, a more realistic topography is required. As we mentioned in the manuscripts, current plume models are still not comprehensive enough. Each model has its own problem of interest and assumptions were made accordingly. Fortunately, we have been seeing efforts on making these models to be more general. As the first trial of using SPH in plume modeling, we prefer to narrow down to the problem of interest and focus on more fundamental aspects.

@@

We have two questions regarding your suggestion: 
1) When we assign pressure value to these static pressure ghost particles, the decrease of pressure due to increase of height ($h$) has already been taken into account. That is to say, pressure of our pressure ghost particles is consistent with meteorological data when establishing the initial condition. Should we still deduce $\rho g h$ from pressure?  
2) Should the total pressure equals to a pressure \textbf{plus} kinetic energy, or minus kinetic energy? Similar questions arise for definition of $p_{\rho}$. Should $\rho g h$ be added to or deduced from pressure? Because total pressure in other fields, such as thermal dynamic analysis of flowing process in engines are defined as pressure plus kinetic energy. %Just to make sure.

In our current model, pressure ghost particles are assumed to be stationary. It does not make a difference whether we use pressure or total pressure since velocity of pressure ghost particles are zero. The next important improvement of our current model under planning is taking the effect of wind field into account. In that case, the pressure ghost particles will not be static any more. The idea of using total pressure can be naturally implemented for that case. Thank you very much for pointing out a potential way for improving pressure outlet boundary condition. \\[6pt]

@@

\textit{deals with the numerical resolution of the equations of the model, not with the agreement between the model and reality. It checks no code errors have been introduced
in the code and one way to do it is a comparison with analytical solutions. I think it is better to simply rename the section ``Applications or Results}

Since JPUE is not the phenomena that we aim to model simulation of that is not validation (validation is to make sure simulations are consistent with reality for the application of interest). Meanwhile, we compared our results against experimental results. We might call such comparison (simulation against measurements) as verification in a more general sense (?).

We added simulation results of shock tube tests which were compared against analytic solutions. By adding such ``pure" verification tests, we keep the title of this section unchanged.
\\[6pt]

@@

\textit{Page 25, line 24. Please look also at the results from this paper: Ezzamel, Adam, Pietro Salizzoni, and Gary R. Hunt. ``Dynamical variability of axisymmetric
buoyant plumes." Journal of Fluid Mechanics 765 (2015): 576-611.}

Thank you for pointing out a source for accessing more recent experimental results. We have added the paper to the references. \citet{ezzamel2015dynamical} did not provide coefficients of their fit equations. So we did not add plots corresponding to  Ezzamel's results in our figures. \\[3pt]
